---
title: "Neuron"
author: "SWK"
date: "02/12/2024"
output: html_document
---

## Set Up
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "IAV-nasal-sc-atlas/02-cell-type-analyses/cell-types")
main.dir = getwd()
dir.create(file.path(main.dir,"neuron"))
```

packages and functions
```{r}
library(tidyverse)
library(Seurat)
library(openxlsx)
library(cowplot)
library(dplyr)
library(future)
theme_set(theme_cowplot())
plan("multicore", workers = 60)
options(future.globals.maxSize = 2048 * 1024^2) #set options for parallelization
source("IAV-nasal-sc-atlas/shared_scripts/plotting_functions_v3.R")
```

Region/Timepoint Palette
```{r}
sample.palette = c("#97F797","#24EA24","#00B700","#007200","#133501",
                   "#C39DEA","#A05CF7","#7E2CFF","#5413BA","#2C0072",
                   "#FFEC64","#F9B903","#F48500","#F24400","#873000")
names(sample.palette) = c("Naive_RM","D02_RM","D05_RM","D08_RM","D14_RM",
                          "Naive_OM","D02_OM","D05_OM","D08_OM","D14_OM",
                          "Naive_LNG","D02_LNG","D02_LNG","D05_LNG","D14_LNG")
```

load in neuron object
```{r}
neuron = readRDS("IAV-nasal-sc-atlas/01-post-cellbender/cell_type_objs/neuron_raw.RDS")
```

We now reprocess the neuron object iteratively until all multiplet/contaminating clusters have been removed

# Round 1
Subset
```{r}
neuron = subset(neuron, features = c(rownames(neuron)[which(rowSums(neuron@assays$SCT@counts > 0) >= 10)], 
                               rownames(neuron[["HTO"]])))
neuron
```

Reprocessing
```{r}
neuron = SCTransform(neuron, verbose = TRUE, method = "glmGamPoi")
neuron = RunPCA(neuron, verbose = TRUE)
DimPlot(neuron, dims = c(1,2), pt.size = 0.1, reduction = "pca", group.by = "orig.ident")
ElbowPlot(neuron, ndims = 50)
neuron = FindNeighbors(neuron, dims = 1:20, k.param = 30, force.recalc = TRUE)
neuron = FindClusters(neuron, resolution = 0.2)
neuron = FindClusters(neuron, resolution = 0.3)
neuron = FindClusters(neuron, resolution = 0.4)
neuron = RunUMAP(neuron, dims = 1:20, n.neighbors = 30)
```

Plot UMAP
```{r}
DimPlot(neuron, pt.size = 0.1, group.by = "SCT_snn_res.0.2", label = TRUE) + NoLegend()
DimPlot(neuron, pt.size = 0.1, group.by = "SCT_snn_res.0.3", label = TRUE) + NoLegend()
#DimPlot(neuron, pt.size = 0.1, group.by = "SCT_snn_res.0.4", label = TRUE) + NoLegend()
DimPlot(neuron, pt.size = 0.1, group.by = "orig.ident", shuffle = TRUE)
DimPlot(neuron, pt.size = 0.1, group.by = "region", shuffle = TRUE)
```

Find Markers
```{r}
neuron = SetIdent(neuron, value = "SCT_snn_res.0.3")
neuron.all.res03.markers = FindAllMarkers(neuron, logfc.threshold = 0.5, min.pct = 0.2, only.pos = TRUE)
write.xlsx(neuron.all.res03.markers, "neuron/neuron_r1_res03_markers.xlsx") #r1 = round 1
```

Vlns/FeaturePlots for contaminating markers
```{r}
VlnPlot(neuron, c("Omp","Epcam","Col3a1","Obp1a","Ptprc","Flt1"), pt.size = -1, stack = TRUE)
```
Cluster 10 is multiplets/garbage
Cluster 12 are dying cells

** Using res = 0.3 **

Now we recluster again!

# Round 2
Remove clusters 10, 12
```{r}
good.neuron.r1 = c(0:9,11,13)
neuron = SetIdent(neuron, value = "SCT_snn_res.0.3")
neuron.r2 = subset(neuron, subset = SCT_snn_res.0.3 %in% good.neuron.r1)
neuron.r2 = subset(neuron.r2, features = c(rownames(neuron.r2)[which(rowSums(neuron.r2@assays$SCT@counts > 0) >= 10)], 
                               rownames(neuron.r2[["HTO"]])))
neuron.r2
```

Reprocess
```{r}
neuron.r2 = SCTransform(neuron.r2, verbose = TRUE, method = "glmGamPoi")
neuron.r2 = RunPCA(neuron.r2, verbose = TRUE)
DimPlot(neuron.r2, dims = c(1,2), pt.size = 0.1, reduction = "pca", group.by = "orig.ident")
ElbowPlot(neuron.r2, ndims = 50)
neuron.r2 = FindNeighbors(neuron.r2, dims = 1:15, k.param = 30, force.recalc = TRUE)
neuron.r2 = FindClusters(neuron.r2, resolution = 0.1)
neuron.r2 = FindClusters(neuron.r2, resolution = 0.2)
#neuron.r2 = FindClusters(neuron.r2, resolution = 0.3)
neuron.r2 = RunUMAP(neuron.r2, dims = 1:15, n.neighbors = 30)
```

Plot UMAP
```{r}
DimPlot(neuron.r2, pt.size = 0.1, group.by = "SCT_snn_res.0.1", label = TRUE)
DimPlot(neuron.r2, pt.size = 0.1, group.by = "SCT_snn_res.0.2", label = TRUE)
#DimPlot(neuron.r2, pt.size = 0.1, group.by = "SCT_snn_res.0.3", label = TRUE)
DimPlot(neuron.r2, pt.size = 0.1, group.by = "orig.ident", shuffle = TRUE)
DimPlot(neuron.r2, pt.size = 0.1, group.by = "region", shuffle = TRUE)
```

Find Markers
```{r}
neuron.r2 = SetIdent(neuron.r2, value = "SCT_snn_res.0.1")
neuron.r2.all.markers.res01 = FindAllMarkers(neuron.r2, logfc.threshold = 0.5, min.pct = 0.2, only.pos = TRUE)
write.xlsx(neuron.r2.all.markers.res01, "neuron/neuron_r2_res01_markers.xlsx", overwrite = TRUE)

neuron.r2 = SetIdent(neuron.r2, value = "SCT_snn_res.0.2")
neuron.r2.all.markers.res02 = FindAllMarkers(neuron.r2, logfc.threshold = 0.5, min.pct = 0.2, only.pos = TRUE)
write.xlsx(neuron.r2.all.markers.res02, "neuron/neuron_r2_res02_markers.xlsx", overwrite = TRUE)

# neuron.r2 = SetIdent(neuron.r2, value = "SCT_snn_res.1")
# neuron.r2.all.markers.res1 = FindAllMarkers(neuron.r2, logfc.threshold = 0.5, min.pct = 0.2, only.pos = TRUE)
# write.xlsx(neuron.r2.all.markers.res1, "neuron/neuron_r2_res1_markers.xlsx", overwrite = TRUE)
```

VlnPlots for contamination
```{r}
neuron.r2 = SetIdent(neuron.r2, value = "SCT_snn_res.0.2")
VlnPlot(neuron.r2, c("Pdlim1","Acsm4","Omp"), pt.size = -1, stack = TRUE) + NoLegend()
```

No clear contaminant clusters!
** USING res = 0.2 **

Assign names from excel sheet
```{r}
neuron.cluster.names = c("OSN","Acsm4+ OSN","Dlg2+ OSN","Differentiating OSN","Cd36+ OSN","Neural Progenitor","Calb2+ OSN","Vomeronasal Sensory Neuron","Emx1+ Stem Cell")
neuron.cluster.names.order = c(5,3,6,0,1,2,4,7,8)+1
neuron.palette = c("#597536","#75c855","#5ec1af","#6d70b2","#8544ba","#ca5b97","#8a413a","#d8593f","#caaa4e")
names(neuron.palette) = neuron.cluster.names[neuron.cluster.names.order]
```
```{r}
neuron.r2$cluster.names = plyr::mapvalues(neuron.r2$SCT_snn_res.0.2, from = 0:8, to = neuron.cluster.names)
neuron.r2$cluster.names = factor(neuron.r2$cluster.names, levels = neuron.cluster.names[neuron.cluster.names.order])
```


# Clustering Results
Sample by cluster
```{r}
neuron.r2.cluster.by.sample = neuron.r2@meta.data %>% group_by(cluster.names,orig.ident) %>% 
  summarise(count = n()) %>% mutate(freq = count/sum(count))
neuron.r2.cluster.by.sample$orig.ident = factor(neuron.r2.cluster.by.sample$orig.ident, levels = names(sample.palette))

ggplot(neuron.r2.cluster.by.sample, aes(x=cluster.names, y=freq*100, fill=orig.ident)) + geom_bar(stat="identity") +
  scale_fill_manual(values = sample.palette) + 
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust = 1)) + labs(x="", y="Proportion of Cluster (%)")
ggsave("neuron/neuron_r2_sample_by_cluster.pdf", height = 5, width = 5)
```


Replot umap with new names
```{r}
DimPlot(neuron.r2, pt.size = 0.1, group.by = "cluster.names", label = TRUE, cols = neuron.r2.palette) + NoLegend()
ggsave("neuron/neuron_r2_named_UMAP.pdf", height = 7, width = 7)
```

Frequency Analysis by cluster and sample
```{r}
neuron.r2$orig.ident = factor(neuron.r2$orig.ident, levels = names(sample.palette)) #some factor fixing for ordering

neuron.RM.freq.by.hash = CalcFreqByHashMD(metadata = neuron.r2@meta.data, 
                                           col = cluster.names, 
                                           which.x.axis = timepoint,
                                           roi = "RM",
                                           palette = sample.palette[1:5], #RM samples
                                           plot.height = 8, plot.width = 10, plot.ncol = 5, 
                                           filename = "neuron/neuron_RM_freq_by_Hash.pdf")
neuron.OM.freq.by.hash = CalcFreqByHashMD(metadata = neuron.r2@meta.data, 
                                           col = cluster.names, 
                                           which.x.axis = timepoint,
                                           roi = "OM",
                                           palette = sample.palette[6:10], #OM samples
                                           plot.height = 8, plot.width = 10, plot.ncol = 5, 
                                           filename = "neuron/neuron_OM_freq_by_Hash.pdf")
neuron.LNG.freq.by.hash = CalcFreqByHashMD(metadata = neuron.r2@meta.data, 
                                            col = cluster.names, 
                                            which.x.axis = timepoint,
                                            roi = "LNG",
                                            palette = sample.palette[11:15], #LNG samples
                                            plot.height = 8, plot.width = 10, plot.ncol = 5, 
                                            filename = "neuron/neuron_LNG_freq_by_Hash.pdf")
```

Big Violin Plot of Cluster Defining Genes
```{r}
neuron.r2 = SetIdent(neuron.r2, value = "cluster.names")

neuron.v3.cluster.genes = c("Omp","Tubb3","Sox11","Neurod1","Gap43","Stmn1","Calb2","Rims3","Stoml3","Nfix",
                           "Acsm4","Nqo1","Dlg2","Kirrel2","Cd36","Cd55","Gnai2","S100z","Emx1","Dmrta2")

VlnPlot(neuron.r2, features = neuron.v3.cluster.genes, pt.size = -1, cols = neuron.r2.palette, stack = TRUE, fill.by = "ident") + NoLegend()
ggsave("neuron/neuron_r2_cluster_markers_violin.pdf", height = 4, width = 8)
```

Save Seurat object
```{r}
saveRDS(neuron.r3, "neuron/neuron_processed.RDS", compress = FALSE)
saveRDS(neuron.r3@meta.data, "neuron/neuron_processed_metadata.RDS", compress = FALSE)
```

