---
title: "Stromal"
author: "SWK"
date: "02/12/2024"
output: html_document
---

** Note, run this script before Endothelial.Rmd, as some endothelial cells snuck into this clustering**

## Set Up
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "IAV-nasal-sc-atlas/02-cell-type-analyses/cell-types")
main.dir = getwd()
dir.create(file.path(main.dir,"stromal"))
```

packages and functions
```{r}
library(tidyverse)
library(Seurat)
library(openxlsx)
library(cowplot)
library(dplyr)
library(future)
theme_set(theme_cowplot())
plan("multicore", workers = 60)
options(future.globals.maxSize = 2048 * 1024^2) #set options for parallelization
source("IAV-nasal-sc-atlas/shared_scripts/plotting_functions_v3.R")
```

Region/Timepoint Palette
```{r}
sample.palette = c("#97F797","#24EA24","#00B700","#007200","#133501",
                   "#C39DEA","#A05CF7","#7E2CFF","#5413BA","#2C0072",
                   "#FFEC64","#F9B903","#F48500","#F24400","#873000")
names(sample.palette) = c("Naive_RM","D02_RM","D05_RM","D08_RM","D14_RM",
                          "Naive_OM","D02_OM","D05_OM","D08_OM","D14_OM",
                          "Naive_LNG","D14_LNG","D02_LNG","D05_LNG","D08_LNG")
```

Load in stromal object
```{r}
stromal = readRDS("IAV-nasal-sc-atlas/01-post-cellbender/cell_type_objs/stromal_raw.RDS")
```

We now reprocess the stromal object iteratively until all multiplet/contaminating clusters have been removed

# Round 1
Remove genes not expressed in this object
```{r}
stromal = subset(stromal, features = c(rownames(stromal)[which(rowSums(stromal@assays$SCT@counts > 0) >= 10)], 
                               rownames(stromal[["HTO"]])))
stromal
```

Reprocessing
```{r}
stromal = SCTransform(stromal, verbose = TRUE, method = "glmGamPoi")
stromal = RunPCA(stromal, verbose = TRUE)
DimPlot(stromal, dims = c(1,2), pt.size = 0.1, reduction = "pca", group.by = "orig.ident")
ElbowPlot(stromal, ndims = 50)
stromal = FindNeighbors(stromal, dims = 1:30, k.param = 30, force.recalc = TRUE)
stromal = FindClusters(stromal, resolution = 0.4)
stromal = FindClusters(stromal, resolution = 0.6)
#stromal = FindClusters(stromal, resolution = 0.8)
stromal = RunUMAP(stromal, dims = 1:30, n.neighbors = 30)
```

Plot UMAP
```{r}
DimPlot(stromal, pt.size = 0.1, group.by = "SCT_snn_res.0.4", label = TRUE) + NoLegend()
DimPlot(stromal, pt.size = 0.1, group.by = "SCT_snn_res.0.6", label = TRUE) + NoLegend()
#DimPlot(stromal, pt.size = 0.1, group.by = "SCT_snn_res.0.8", label = TRUE) + NoLegend()
DimPlot(stromal, pt.size = 0.1, group.by = "orig.ident", shuffle = TRUE)
DimPlot(stromal, pt.size = 0.1, group.by = "region", shuffle = TRUE)
```

Find Markers
```{r}
# stromal = SetIdent(stromal, value = "SCT_snn_res.0.4")
# stroma.all.res04.markers = FindAllMarkers(stromal, logfc.threshold = 0.5, min.pct = 0.2, only.pos = TRUE)
# write.xlsx(stroma.all.res04.markers, "stroma_r1_res04_markers.xlsx") #r1 = round 1

stromal = SetIdent(stromal, value = "SCT_snn_res.0.6")
stromal.all.res06.markers = FindAllMarkers(stromal, logfc.threshold = 0.5, min.pct = 0.2, only.pos = TRUE)
write.xlsx(stromal.all.res06.markers, "stromal/stromal_r1_res06_markers.xlsx") #r1 = round 1
```

Vlns/FeaturePlots for contaminating markers
```{r}
VlnPlot(stromal, c("Omp","Epcam","Pecam1","Ptprc","Obp1a"), pt.size = -1, stack = TRUE)
```
Cluster 1, 4, 5 have OMP/Epcam
Cluster 7 are endothelial cells (remove and keep for later)
Cluster 12 is macrophage doublets
Cluster 13, 14 are epithelial doublets

** Using res = 0.6 **

Save cluster 7 separately
```{r}
stromal.endo = subset(stromal, subset = SCT_snn_res.0.6 == 7)
saveRDS(stromal.endo, file = "stromal/stromal_endothelial.RDS", compress = FALSE)
```

Now we recluster again!

# Round 2
Remove clusters 4, 5, 7, 8, 11-13
```{r}
good.stroma.r1 = c(0,2,3,6,8:11) #cluster to keep
stromal = SetIdent(stromal, value = "SCT_snn_res.0.6")
stromal.r2 = subset(stromal, subset = SCT_snn_res.0.6 %in% good.stroma.r1)
stromal.r2 = subset(stromal.r2, features = c(rownames(stromal.r2)[which(rowSums(stromal.r2@assays$SCT@counts > 0) >= 10)], 
                               rownames(stromal.r2[["HTO"]])))
stromal.r2
```

Reprocess
```{r}
stromal.r2 = SCTransform(stromal.r2, verbose = TRUE, method = "glmGamPoi")
stromal.r2 = RunPCA(stromal.r2, verbose = TRUE)
DimPlot(stromal.r2, dims = c(1,2), pt.size = 0.1, reduction = "pca", group.by = "orig.ident")
ElbowPlot(stromal.r2, ndims = 50)
stromal.r2 = FindNeighbors(stromal.r2, dims = 1:15, k.param = 30, force.recalc = TRUE)
stromal.r2 = FindClusters(stromal.r2, resolution = 0.4)
stromal.r2 = FindClusters(stromal.r2, resolution = 0.6)
stromal.r2 = RunUMAP(stromal.r2, dims = 1:15, n.neighbors = 30)
```

Plot UMAP
```{r}
DimPlot(stromal.r2, pt.size = 0.1, group.by = "SCT_snn_res.0.4", label = TRUE) + NoLegend()
DimPlot(stromal.r2, pt.size = 0.1, group.by = "SCT_snn_res.0.6", label = TRUE) + NoLegend()
DimPlot(stromal.r2, pt.size = 0.1, group.by = "orig.ident", shuffle = TRUE)
DimPlot(stromal.r2, pt.size = 0.1, group.by = "region", shuffle = TRUE)
```

Find Markers
```{r}
stromal.r2 = SetIdent(stromal.r2, value = "SCT_snn_res.0.4")
stromal.r2.all.markers.res04 = FindAllMarkers(stromal.r2, logfc.threshold = 0.5, min.pct = 0.2, only.pos = TRUE)
write.xlsx(stromal.r2.all.markers.res04, "stromal/stromal_r2_res04_markers.xlsx", overwrite = TRUE)

stromal.r2 = SetIdent(stromal.r2, value = "SCT_snn_res.0.6")
stromal.r2.all.markers.res06 = FindAllMarkers(stromal.r2, logfc.threshold = 0.5, min.pct = 0.2, only.pos = TRUE)
write.xlsx(stromal.r2.all.markers.res06, "stromal/stromal_r2_res06_markers.xlsx", overwrite = TRUE)
```

Cluster 10 are multiplets on res = 0.6
** Choosing res = 0.6**


# Round 3
Remove final multiplet cluster 10 using res = 0.6
```{r}
good.stroma.r2 = c(0:9)
stromal.r3 = subset(stromal.r2, subset = SCT_snn_res.0.6 %in% good.stroma.r2)
stromal.r3 = subset(stromal.r3, features = c(rownames(stromal.r3)[which(rowSums(stromal.r3@assays$SCT@counts > 0) >= 10)], 
                               rownames(stromal.r3[["HTO"]])))
stromal.r3
```

Reprocess
```{r}
stromal.r3 = SCTransform(stromal.r3, verbose = TRUE, method = "glmGamPoi")
stromal.r3 = RunPCA(stromal.r3, verbose = TRUE)
DimPlot(stromal.r3, dims = c(1,2), pt.size = 0.1, reduction = "pca", group.by = "orig.ident")
ElbowPlot(stromal.r3, ndims = 50)
stromal.r3 = FindNeighbors(stromal.r3, dims = 1:15, k.param = 30)
stromal.r3 = FindClusters(stromal.r3, resolution = 0.3)
stromal.r3 = FindClusters(stromal.r3, resolution = 0.4)
stromal.r3 = FindClusters(stromal.r3, resolution = 0.5)
stromal.r3 = RunUMAP(stromal.r3, dims = 1:15, n.neighbors = 30)
```

Plot UMAP
```{r}
DimPlot(stromal.r3, pt.size = 0.1, group.by = "SCT_snn_res.0.3", label = TRUE) + NoLegend()
DimPlot(stromal.r3, pt.size = 0.1, group.by = "SCT_snn_res.0.4", label = TRUE) + NoLegend()
DimPlot(stromal.r3, pt.size = 0.1, group.by = "SCT_snn_res.0.5", label = TRUE) + NoLegend()
DimPlot(stromal.r3, pt.size = 0.1, group.by = "orig.ident", shuffle = TRUE)
DimPlot(stromal.r3, pt.size = 0.1, group.by = "region", shuffle = TRUE)
```

Find Markers
```{r}
stromal.r3 = SetIdent(stromal.r3, value = "SCT_snn_res.0.3")
stromal.r3.all.markers.res03 = FindAllMarkers(stromal.r3, logfc.threshold = 0.5, min.pct = 0.2, only.pos = TRUE)
write.xlsx(stromal.r3.all.markers.res03, "stromal/stromal_r3_res03_markers.xlsx", overwrite = TRUE)

# stromal.r3 = SetIdent(stromal.r3, value = "SCT_snn_res.0.4")
# stromal.r3.all.markers.res04 = FindAllMarkers(stromal.r3, logfc.threshold = 0.5, min.pct = 0.2, only.pos = TRUE)
# write.xlsx(stromal.r3.all.markers.res04, "stromal/stromal_r3_res04_markers.xlsx", overwrite = TRUE)
```

** We choose res = 0.3 **

We then annotated clusters based on marker expression from the output above and name them in the object
```{r}
stromal.cluster.names = c("SMC","Pericyte 1","Schwann 1","Pericyte 2",
                         "Schwann 2","Oligodendrocyte","OBP SMC/Pericyte","Pcp4+Nrtn+ Stroma")
stromal.cluster.names.order = c(0,1,3,6,7,2,4,5)+1 #we order for plotting
stromal.palette = c("#83d55b","#1389a7","#6357b1","#59d2b7","#ceb44b","#bc4573","#bf57c7","#bf5837")
names(stromal.palette) = stromal.cluster.names[stromal.cluster.names.order]
```
```{r}
stromal.r3$cluster.names = plyr::mapvalues(stromal.r3$SCT_snn_res.0.3, from = 0:7, to = stromal.cluster.names)
stromal.r3$cluster.names = factor(stromal.r3$cluster.names, levels = stromal.cluster.names[stromal.cluster.names.order])
```


# Clustering Results
Sample by cluster breakdown
```{r}
stromal.r3.cluster.by.sample = stromal.r3@meta.data %>% group_by(cluster.names,orig.ident) %>% 
  summarise(count = n()) %>% mutate(freq = count/sum(count))
stromal.r3.cluster.by.sample$orig.ident = factor(stromal.r3.cluster.by.sample$orig.ident, levels = names(sample.palette))

ggplot(stromal.r3.cluster.by.sample, aes(x=cluster.names, y=freq*100, fill=orig.ident)) + geom_bar(stat="identity") +
  scale_fill_manual(values = sample.palette) + 
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust = 1)) + labs(x="", y="Proportion of Cluster (%)")
ggsave("stromal/stromal_sample_by_cluster.pdf", height = 5, width = 4.5)
```

Replot UMAP with new names
```{r}
DimPlot(stromal.r3, pt.size = 0.5, group.by = "cluster.names", label = TRUE, cols = stromal.palette) + NoLegend()
ggsave("stromal/stromal_named_UMAP.pdf", height = 6, width = 6)
```

Frequency Analysis by cluster and sample per region
```{r}
stromal.r3$orig.ident = factor(stromal.r3$orig.ident, levels = names(sample.palette)) #some factor fixing for ordering

stromal.RM.freq.by.hash = CalcFreqByHashMD(metadata = stromal.r3@meta.data, 
                                           col = cluster.names, 
                                           which.x.axis = timepoint,
                                           roi = "RM",
                                           palette = sample.palette[1:5], #RM samples
                                           plot.height = 8, plot.width = 10, plot.ncol = 5, 
                                           filename = "stromal/stromal_RM_freq_by_Hash.pdf")
stromal.OM.freq.by.hash = CalcFreqByHashMD(metadata = stromal.r3@meta.data, 
                                           col = cluster.names, 
                                           which.x.axis = timepoint,
                                           roi = "OM",
                                           palette = sample.palette[6:10], #OM samples
                                           plot.height = 8, plot.width = 10, plot.ncol = 5, 
                                           filename = "stromal/stromal_OM_freq_by_Hash.pdf")
stromal.LNG.freq.by.hash = CalcFreqByHashMD(metadata = stromal.r3@meta.data, 
                                            col = cluster.names, 
                                            which.x.axis = timepoint,
                                            roi = "LNG",
                                            palette = sample.palette[11:15], #LNG samples
                                            plot.height = 8, plot.width = 10, plot.ncol = 5, 
                                            filename = "stromal/stromal_LNG_freq_by_Hash.pdf")
```

Big Violin Plot of Cluster Defining Genes
```{r}
stromal.r3 = SetIdent(stromal.r3, value = "cluster.names")

stromal.v3.cluster.genes = c("Acta2","Tagln","Kcnj8","Pdgfrb","Fmod","Ccl19","Obp1b","Pcp4","Nrtn","Corin",
                             "Fabp7","Ncam1","Cldn5","Scn7a","Gfra3","Slitrk6","Mbp","Mpz","Ncmap")

VlnPlot(stromal.r3, features = stromal.v3.cluster.genes, pt.size = -1, 
        cols = stroma.r3.palette, stack = TRUE, fill.by = "ident") + NoLegend()
ggsave("stromal/stromal_cluster_markers_violin.pdf", height = 4, width = 8)
```

Save Seurat object
```{r}
saveRDS(stromal.r3, "stromal/stromal_processed.RDS", compress = FALSE)
saveRDS(stromal.r3@meta.data, "stromal/stromal_processed_metadata.RDS", compress = FALSE)
```