---
title: "create_sobj"
author: "SWK"
date: "02/07/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "IAV-nasal-sc-atlas/01-post-cellbender/")
main.dir = getwd()
dir.create(file.path(main.dir,"output"))
```

```{r}
library(tidyverse)
library(Seurat)
library(cowplot)
library(future)
theme_set(theme_cowplot())
#plan("multicore", workers = 60)
#options(future.globals.maxSize = 2048 * 1024^2) #set options for parallelization
```

# Importing Data
## Cellbender h5 files
Read in each cellbender filtered matrix and fix names
```{r}
#list all of the files so we can read them in recursively
h5.cellbender.files = list.files("inputs/cb", pattern = "h5", full.names = TRUE)

#name order of the files
sample.names = c("D14_OM","D14_RM","D14_LNG","D02_OM","D02_RM","D02_LNG","D05_OM","D05_RM","D05_LNG",
                 "D08_OM","D08_RM","D08_LNG","Naive_OM","Naive_RM","Naive_LNG")

#create an empty list to hold the matrices
cb.mats = list()

#loop through all of the files
for (i in 1:length(h5.cellbender.files)){
  cb.mats[[i]] = Read10X_h5(h5.cellbender.files[i])
}

#remove "-1" from all of the column names and order, and add sample name on the top
for (i in 1:length(cb.mats)){
  colnames(cb.mats[[i]]) = sapply(strsplit(colnames(cb.mats[[i]]), split = "-"), "[[", 1)
  colnames(cb.mats[[i]]) = paste(sample.names[i], colnames(cb.mats[[i]]), sep=".")
  cb.mats[[i]] = cb.mats[[i]][,order(colnames(cb.mats[[i]]))]
}
```

What are the cell numbers in each?
```{r}
sapply(cb.mats, function(x) dim(x)[2])
```


# Prefiltering
## Merge Matrices and create Seurat Object
Merge matrices
```{r}
mat.big = cbind(cb.mats[[1]], cb.mats[[2]])
cb.mats = cb.mats[3:15]
times = length(cb.mats)-1

for(i in 1:times){
  mat.big = cbind(mat.big, cb.mats[[1]])
  cb.mats = cb.mats[-1]
}

mat.big = cbind(mat.big, cb.mats[[1]])
rm(cb.mats)
```

Create a big seurat object
```{r}
cb = CreateSeuratObject(counts = mat.big, names.field = 1, names.delim = "\\.", min.cells = 0, min.features = 0)
cb
rm(mat.big)
table(cb$orig.ident)
```


## Add in HTO data
The calls and hash counts were pulled from the cumulus demux pipeline and are provided in the git repo
```{r}
load("data/EM_demux_primary.RData")
```

Trim demux down to bcs in mat.big and add in empty rows for missing cells
```{r}
#rename
for(i in 1:length(demux.calls)){
  row.names(demux.calls[[i]]) = paste(sample.names[i], row.names(demux.calls[[i]]), sep = ".")
  row.names(demux.counts[[i]]) = paste(sample.names[i], row.names(demux.counts[[i]]), sep = ".")
}

#keep only bcs present in the seurat object
for(i in 1:length(demux.calls)){
  demux.calls[[i]] = demux.calls[[i]][which(rownames(demux.calls[[i]]) %in% colnames(cb)),]
  demux.counts[[i]] = as.data.frame(demux.counts[[i]][which(rownames(demux.counts[[i]]) %in% colnames(cb)),])
  demux.counts[[i]]$cell.name = row.names(demux.counts[[i]])
}

#some clean up
demux.calls.trim = do.call(rbind, demux.calls)
demux.counts.trim = demux.counts %>% reduce(full_join, by = "cell.name")
row.names(demux.counts.trim) = demux.counts.trim$cell.name
demux.counts.trim = demux.counts.trim[,-4]
demux.counts.trim[is.na(demux.counts.trim)] = 0

#additional rows for missing cells in demux counts
missing.cells = colnames(cb)[which(! colnames(cb) %in% rownames(demux.counts.trim))]
missing.cells.counts = as.data.frame(matrix(data = 0, nrow = length(missing.cells), ncol = ncol(demux.counts.trim)))
rownames(missing.cells.counts) = missing.cells; colnames(missing.cells.counts) = colnames(demux.counts.trim)
demux.counts.trim = rbind(demux.counts.trim, missing.cells.counts)

#order everything
demux.counts.trim = demux.counts.trim[order(rownames(demux.counts.trim)),]
demux.calls.trim = demux.calls.trim[order(rownames(demux.calls.trim)),]
```

Add in the HTOs and calls to cb
```{r}
cb = AddMetaData(cb, metadata = demux.calls.trim[,c("demux_type","assignment")])
cb[["HTO"]] = CreateAssayObject(counts = t(demux.counts.trim))
```

remove unneccesary demux stuff now
```{r}
rm(demux.calls, demux.calls.trim, demux.counts, demux.counts.trim, missing.cells.counts)
gc()
```


## QC Filtering
MT read percentage
```{r}
cb <- PercentageFeatureSet(cb, pattern = "^mt-", col.name = "percent.mt")
```

VlnPlots of QC metrics
```{r}
VlnPlot(cb, features = c("nCount_RNA", "nFeature_RNA", "percent.mt"), pt.size = 0, ncol = 1, group.by = "orig.ident")
ggsave("output/QC Violins.pdf", height = 10, width = 7)
```

Remove bad cells
Here we remove any cells with fewer than 500 genes, fewer than 750 or more than 100000 UMIs, and more than 15% mitochondrial reads
```{r}
cb = subset(cb, subset = nFeature_RNA > 500 & percent.mt < 15 & 
              nCount_RNA > 750 & nCount_RNA < 100000)
```

Remove any cells with more than 10000 hashes and any genes not expressed in at least 10 cells
```{r}
cb = subset(cb, subset = nCount_HTO <= 1e4)
cb = subset(cb, features = c(rownames(cb)[which(rowSums(cb@assays$RNA@counts > 0) >= 10)], rownames(cb[["HTO"]])))
```

QC Metric plots after filtering
```{r}
VlnPlot(cb, features = c("nCount_RNA", "nFeature_RNA", "percent.mt"), pt.size = 0, ncol = 1, group.by = "orig.ident")
ggsave("output/QC Violins Trimmed.pdf", height = 10, width = 7.5)
```

Plot the Cell Counts per sample
```{r}
cell.counts.trimmed = data.frame(table(cb$orig.ident))
ggplot(cell.counts.trimmed, aes(x=Var1, y=Freq)) + geom_bar(aes(fill = sample.names), stat = "identity") + labs(x = "", y = "cb Cell Count") + guides(fill = 'none') + theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggsave("output/Cell Counts Trimmed v3.pdf", height = 5, width = 7)
```

Output combined data
```{r}
saveRDS(cb, file = "output/cellbender_trimmed_sobj.RDS", compress = FALSE)
```

