---
title: "clustering_and_filtering"
author: "SWK"
date: "02/10/2024"
output: html_document
---

# Set Up
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "IAV-nasal-sc-atlas/01-post-cellbender")
```

Note, this analysis was done using Seurat v4.2.0
```{r}
library(tidyverse)
library(Seurat)
library(openxlsx)
library(cowplot)
library(dplyr)
library(future)
theme_set(theme_cowplot())
#plan("multicore", workers = 60)
#options(future.globals.maxSize = 2048 * 1024^2) #set options for parallelization

sample.names = c("D14_OM","D14_RM","D14_LNG","D02_OM","D02_RM","D02_LNG","D05_OM","D05_RM","D05_LNG",
                 "D08_OM","D08_RM","D08_LNG","Naive_OM","Naive_RM","Naive_LNG")

main.dir = getwd()
```

## Bring in data
Load in Seurat Object
```{r}
load("cellbender Seurat Trimmed v3.RData")
```

# Top Level Clustering and Exploration, doublets removed
Create folder for outputs
```{r}
dir.create(file.path(main.dir,"dedub"))
```

## remove doublets
```{r}
cb.dedub = subset(cb, cells = colnames(cb)[cb$demux_type %in% c("unknown","singlet")])
cb.dedub
rm(cb)
gc()
```

## processing through clustering
Run scTransform
```{r}
cb.dedub = SCTransform(cb.dedub, verbose = TRUE, method = "glmGamPoi")
gc()
```

Run PCA, look at plenty of PCs on the elbow plot since the dataset is quite complex
```{r}
cb.dedub = RunPCA(cb.dedub, verbose = TRUE)
DimPlot(cb.dedub, dims = c(1,2), pt.size = 0.1)
ElbowPlot(cb.dedub, ndims = 75)
```

We picked 40 PCs on the Elbow Plot for downstream analysis.
Generally, we try a few different clustering resolutions to help determine over/underclustering
```{r}
cb.dedub = FindNeighbors(cb.dedub, dims = 1:40, k.param = 30)
cb.dedub = FindClusters(cb.dedub, dims = 1:40, resolution = 0.2)
cb.dedub = FindClusters(cb.dedub, dims = 1:40, resolution = 0.4)
cb.dedub = FindClusters(cb.dedub, dims = 1:40, resolution = 0.6)
cb.dedub = RunUMAP(cb.dedub, dims = 1:40, n.neighbors = 30)
```

Plot UMAP with various clustering resolutions, sample identity, demultiplex type, and UMI count
```{r}
p1 = DimPlot(cb.dedub, dims = 1:2, pt.size = 0.3, group.by = "SCT_snn_res.0.2", raster = TRUE)
p2 = DimPlot(cb.dedub, dims = 1:2, pt.size = 0.3, group.by = "SCT_snn_res.0.4", raster = TRUE)
p3 = DimPlot(cb.dedub, dims = 1:2, pt.size = 0.3, group.by = "SCT_snn_res.0.6", raster = TRUE)
p4 = DimPlot(cb.dedub, dims = 1:2, pt.size = 0.3, group.by = "orig.ident", raster = TRUE)
p5 = DimPlot(cb.dedub, dims = 1:2, pt.size = 0.3, group.by = "demux_type", raster = TRUE)
p6 = FeaturePlot(cb.dedub, features = "nCount_SCT", pt.size = 0.3, raster = TRUE)

plot_grid(p1,p2,p3,p4,p5,p6, ncol = 3)
ggsave("dedub_v1_UMAP.pdf", height = 12, width = 20)
```

Find cluster marker genes
We choose res = 0.6 in order to identify clusters of doublets/low quality cells for removal
```{r}
#cb.dedub = SetIdent(cb.dedub, value = "SCT_snn_res.0.2")
#all.markers.res02 = FindAllMarkers(cb.dedub, logfc.threshold = 0.5, min.pct = 0.2, only.pos = TRUE, max.cells.per.ident = 1500, densify = TRUE)
#cb.dedub = SetIdent(cb.dedub, value = "SCT_snn_res.0.4")
#all.markers.res04 = FindAllMarkers(cb.dedub, logfc.threshold = 0.5, min.pct = 0.2, only.pos = TRUE, max.cells.per.ident = 1500)
cb.dedub = SetIdent(cb.dedub, value = "SCT_snn_res.0.6")
all.markers.res06 = FindAllMarkers(cb.dedub, logfc.threshold = 0.5, min.pct = 0.2, only.pos = TRUE, max.cells.per.ident = 1500)
```

Write out the marker gene table
```{r}
#write.xlsx(all.markers.res02, file = "dedub/scT_dedub_res02_all_markers.xlsx")
#write.xlsx(all.markers.res04, file = "dedub/scT_dedub_res04_all_markers.xlsx")
write.xlsx(all.markers.res06, file = "dedub/scT_dedub_res06_all_markers.xlsx")
```

Plotting for visual inspection purposes
We look at lineage and spillover genes to help call multiplets/bad clusters
```{r}
DimPlot(cb.dedub, dims = 1:2, pt.size = 0.3, group.by = "SCT_snn_res.0.6", 
        raster = TRUE, label = TRUE, label.size = 4, label.box = TRUE)
VlnPlot(cb.dedub, c("Omp","Epcam","Obp1a","Obp2a","Ptprc","Sox11","Col1a1","Pdpn","Pecam1"), 
        pt.size = -1, stack = TRUE, fill.by = "ident", flip = TRUE) + NoLegend()
```

## Filter out bad clusters
We are going to remove the following clusters that
1) express multiple mutually exclusive lineage genes
2) only express very high levels of OBP genes (likely exogenous RNA overwhelming true cell identity) 
3) lack any known lineage markers

Thus, we remove clusters: 1, 12, 17, 20, 30, 31, 46
```{r}
good.clusters = c(0,2:11,13:16,18,19,21:29,32:45,47,48)
cb.dedub = SetIdent(cb.dedub, value = "SCT_snn_res.0.6")
cb.tr1 = subset(cb.dedub, subset = SCT_snn_res.0.6 %in% good.clusters)
cb.tr1
gc()
#rm(cb.dedub) #to save space you can save and/or remove the prior object
```

We will now repeat the processing pipeline procedure to find finer grained multiplets/bad cells

# Top Level Clustering and Exploration, trim round 1 (tr1)
## Process through once more
Make new folder for outputs
```{r}
dir.create(file.path(main.dir,"tr1"))
```

scTransform
```{r}
cb.tr1 = SCTransform(cb.tr1, verbose = TRUE, method = "glmGamPoi")
gc()
```

PCA and clustering (still using 40 PCs)
```{r}
cb.tr1 = RunPCA(cb.tr1, verbose = TRUE)
DimPlot(cb.tr1, dims = c(1,2), pt.size = 0.1, reduction = "pca", group.by = "orig.ident")
ElbowPlot(cb.tr1, ndims = 50)
cb.tr1 = FindNeighbors(cb.tr1, dims = 1:40, k.param = 30)
cb.tr1 = FindClusters(cb.tr1, resolution = 0.6)
cb.tr1 = RunUMAP(cb.tr1, dims = 1:40, n.neighbors = 30)
```

Plot UMAPs
```{r}
p1 = DimPlot(cb.tr1, dims = 1:2, pt.size = 0.1, group.by = "SCT_snn_res.0.6")
p3 = DimPlot(cb.tr1, dims = 1:2, pt.size = 0.1, group.by = "orig.ident")
p4 = DimPlot(cb.tr1, dims = 1:2, pt.size = 0.1, group.by = "demux_type")

plot_grid(p1,p2,p3, ncol = 2)
ggsave("tr1/tr1_UMAP_res06.pdf", height = 16, width = 16)
```

Generate marker list and save output
```{r}
all.markers.dedub.t1.res06 = FindAllMarkers(cb.tr1, logfc.threshold = 0.5, min.pct = 0.2, only.pos = TRUE, max.cells.per.ident = 1500)
write.xlsx(all.markers.dedub.t1.res06, "dedub trim1/cbdedubt1_sct_res06_all_markers.xlsx")
```

Visual inspection like before
```{r}
DimPlot(cb.tr1, dims = 1:2, pt.size = 0.3, group.by = "SCT_snn_res.0.6", 
        raster = TRUE, label = TRUE, label.size = 4, label.box = TRUE)
VlnPlot(cb.tr1, c("Omp","Epcam","Obp1a","Obp2a","Ptprc","Sox11","Col3a1","Pdpn","Pecam1"), 
        pt.size = -1, stack = TRUE, fill.by = "ident", flip = TRUE) + NoLegend()
```

We look at some specific comparison just to understand if these are subsets/states or doublets
```{r}
markers.3v31 = FindMarkers(cb.tr1, ident.1 = 3, ident.2 = 31, logfc.threshold = 0.5, min.pct = 0.2)
markers.34 = FindMarkers(cb.tr1, ident.1 = 34, logfc.threshold = 0.5, min.pct =TRUE, only.pos = TRUE, max.cells.per.ident = 1500)
markers.16v41 = FindMarkers(cb.tr1, ident.1 = 16, ident.2 = 41, logfc.threshold = 0.5, min.pct = 0.2)
```

Finally, there are two clusters that seem to have snuck through as doublets
```{r}
good.clusters2 = c(0:35,38:43)
cb.tr2 = subset(cb.tr1, subset = SCT_snn_res.0.6 %in% good.clusters2)
```

## Assigning Broad Cell Type Classes
Based on the lineage genes from the violin plot above, as well as proximity on the UMAP, 
and knowledge of other lineage genes, we assign a cell type identity to each cluster.
We also add in metadata columns for timepoint and region
Note: we have a few clusters defined by cell cycle genes (e.g., Mki67). We annotate these and return to them below.
```{r}
#cell.type (from DE gene xlsx sheet)
cell.types = c("Neuron","Myeloid","Neuron","Epithelial","Neuron","T/NK Cell","Myeloid",
               "B Cell","Endothelial","Neuron","Granulocyte","Neuron","Fibroblast","Myeloid",
               "Epithelial","Epithelial","Epithelial","Stromal","Fibroblast","Epithelial",
               "Epithelial","Granulocyte","Epithelial","Granulocyte","B Cell",
               "Epithelial","Epithelial","Endothelial","Cycling","Stromal","Epithelial",
               "Neuron","Epithelial","Cycling","Epithelial","Epithelial","Fibroblast",
               "Neuron","Epithelial","Epithelial","Epithelial","T/NK Cell")
cb.tr2$cell.type = plyr::mapvalues(cb.tr2$SCT_snn_res.0.6, from = c(0:35,38:43), to = cell.types)

#Time point
cb.tr2$timepoint = sapply(strsplit(x = as.character(cb.tr2$orig.ident), split = "_"), "[[",1)

#region
cb.tr2$region = sapply(strsplit(x = as.character(cb.tr2$orig.ident), split = "_"), "[[",2)
```

create directory for trim round 2
```{r}
dir.create(file.path(main.dir,"tr2"))
```

## Annotating cycling/proliferating cells
We subset the clusters labeled as "Cycling" by cell.type to assign cell type identities
```{r}
cb.tr2 = SetIdent(cb.tr2, value = "cell.type")
cb.prolif = subset(cb.tr2, idents = "Cycling")
cb.prolif = subset(cb.prolif, features = c(rownames(cb.prolif)[which(rowSums(cb.prolif@assays$SCT@counts > 0) >= 10)], 
                               rownames(cb.prolif[["HTO"]])))
cb.prolif
```

Reprocess again. Here we choose 10 PCs
```{r}
cb.prolif = SCTransform(cb.prolif, verbose = TRUE, method = "glmGamPoi")
cb.prolif = RunPCA(cb.prolif, verbose = TRUE)
DimPlot(cb.prolif, dims = c(1,2), pt.size = 0.1, reduction = "pca", group.by = "orig.ident")
ElbowPlot(cb.prolif, ndims = 50)
cb.prolif = FindNeighbors(cb.prolif, dims = 1:10, k.param = 30)
cb.prolif = FindClusters(cb.prolif, resolution = 0.2)
cb.prolif = RunUMAP(cb.prolif, dims = 1:10, n.neighbors = 30)
```

Plot UMAP
```{r}
DimPlot(cb.prolif, pt.size = 0.1, group.by = "SCT_snn_res.0.2", label = TRUE) + NoLegend()
DimPlot(cb.prolif, pt.size = 0.1, group.by = "orig.ident") + NoLegend()
```

Find Markers and write out results
```{r}
cb.prolif = SetIdent(cb.prolif, value = "SCT_snn_res.0.2")
prolif.all.markers = FindAllMarkers(cb.prolif, logfc.threshold = 0.5, min.pct = 0.2, only.pos = TRUE)
View(prolif.all.markers)
write.xlsx(prolif.all.markers, "tr2/prolif_res02_all_markers.xlsx")
```

Visualize markers
```{r}
FeaturePlot(cb.prolif, features = c("Omp","Epcam","Ptprc","Col3a1","Cd3g"), order = TRUE) & NoLegend() & NoAxes()
VlnPlot(cb.prolif, features = c("Omp","Epcam","Ptprc","Col3a1"), pt.size = -1, stack = TRUE)
```

We quickly annotate:
0 = low quality cells; 1 = B Cell; 2 = Monocyte; 3 = HSC; 4 = Epithelial; 
5 = Neuron; 6 = Granulocyte, 7 = B cell; 8 = Mast cell; 9 = B Cell

```{r}
cb.prolif$cell.type = plyr::mapvalues(cb.prolif$seurat_clusters, from=c(0:9), 
                            to=c("Multiplet","B Cell", "Myeloid", "HSC", "Epithelial", 
                                 "Neuron", "Granulocyte", "B Cell", "Granulocyte", "B Cell"))
```

We then replace the "Cycling" name with the newly identified cell types and remove the multiplets
```{r}
cb.tr2$cell.type[cb.tr2$cell.type == "Cycling"] = as.character(cb.prolif$cell.type)
cb.tr2 = subset(cb.tr2, subset = cell.type %in% c("Neuron","Myeloid","Epithelial","T/NK Cell", "B Cell", 
                                           "Endothelial", "Granulocyte", "Fibroblast", "Stromal", "HSC"))
gc()
```

Save the filtered object for later use
```{r}
saveRDS(cb.tr2, file = "tr2/full_dataset_post_filter.RDS", compress = FALSE) #can compress, but takes quite a long time to save
```

# Split up into cell type objects for downstream annotation
For annotation of cells with PR8 reads, we quickly add in a column to the metadata that counts the total number of PR8 UMIs per cell
```{r}
cb.tr2$PR8.SCT.total = colSums(cb.tr2@assays$SCT@data[grep("PR8", rownames(cb.tr2), value = TRUE),])
```

Create output directory for cell type objects
```{r}
dir.create(file.path(main.dir,"cell_type_objs"))
```

We simply split our object into one for each cell type (minus the HSCs) and save each one for further processing
```{r}
cb.tr2 = SetIdent(cb.tr2, value = "cell.type")
neuron = subset(cb.tr2, cell.type == "Neuron")
myeloid = subset(cb.tr2, cell.type == "Myeloid")
epi = subset(cb.tr2, cell.type == "Epithelial")
TNK = subset(cb.tr2, cell.type == "T/NK Cell")
B = subset(cb.tr2, cell.type == "B Cell")
endo = subset(cb.tr2, cell.type == "Endothelial")
granulo = subset(cb.tr2, cell.type == "Granulocyte")
fibro = subset(cb.tr2, cell.type == "Fibroblast")
stromal = subset(cb.tr2, cell.type == "Stromal")
```

save the objects
```{r}
saveRDS(neuron, file = "cell_type_objs/neuron_raw.RDS", compress = FALSE)
saveRDS(myeloid, file = "cell_type_objs/myeloid_raw.RDS", compress = FALSE)
saveRDS(epi, file = "cell_type_objs/epithelial_raw.RDS", compress = FALSE)
saveRDS(TNK, file = "cell_type_objs/TNK_raw.RDS", compress = FALSE)
saveRDS(B, file = "cell_type_objs/Bcell_raw.RDS", compress = FALSE)
saveRDS(endo, file = "cell_type_objs/endothelial_raw.RDS", compress = FALSE)
saveRDS(granulo, file = "cell_type_objs/granulocyte_raw.RDS", compress = FALSE)
saveRDS(fibro, file = "cell_type_objs/fibroblast_raw.RDS", compress = FALSE)
saveRDS(stromal, file = "cell_type_objs/stromal_raw.RDS", compress = FALSE)
```

Next steps are in "02-cell-type-analyses"
Full dataset analysis (03-full-data-set) will be performed after each of the cell type analyses are complete!
