---
title: "rechallenge_analysis"
author: "SWK"
date: '2024-02-28'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "IAV-nasal-sc-atlas/04-rechallenge")
```

packages
```{r}
library(tidyverse)
library(Seurat)
library(cowplot)
library(openxlsx)
library(future)
library(propr)
theme_set(theme_cowplot())
# plan("multicore", workers = 30)
# options(future.globals.maxSize = 2000 * 1024^2) #set options for parallelization
source("IAV-nasal-sc-atlas/shared_scripts/plotting_functions_v3.R")
```

load in data
```{r}
RM = readRDS(file = "IAV-nasal-sc-atlas/03-full-data-set/data-out/RM_annotated.RDS")
rechallenge = readRDS(file = "output/rechallenge_annotated.RDS")
```

New palette
```{r}
RM.palette.all = c("#97F797","#24EA24","#00B700","#007200","#133501","#081402","#03befc","#0065b8","#757575","#ff5ef4","#b8339b")
names(RM.palette.all) = c("Naive","D02","D05","D08","D14","D60_PR8","C2_PR8","C5_PR8","D60_X31","C2_X31","C5_X31")

cell.type.palette = c("#cb9b92","#c15336","#cdbc52","#546439","#79ce5e","#82c7b7","#8089c0","#8447c0","#4d2f46","#c85191")
names(cell.type.palette) = c("Neuron","Epithelial","Myeloid","Granulocyte","B Cell","T/NK Cell","Endothelial","Fibroblast","Stromal","HSC")
```

# Compositional Analysis
## Projection into primary infection trajectory
Bring in the precalculated compositional data
```{r}
#full.freq.table = readRDS(file = "data-out/full_dataset_freq_table.RDS")
CP3k.mat = readRDS(file = "data-out/full_dataset_CP3k_mat.RDS")
label.metadata = readRDS(file = "data-out/full_dataset_comp_metadata.RDS")
```

Re-calculate the RM compositional space from the primary infection timecourse
```{r}
# scale
CP3k.mat.RM = CP3k.mat[,label.metadata$region == "RM"]
CP3k.mat.RM = CP3k.mat.RM[rowSums(CP3k.mat.RM) > 0,]
label.metadata.RM = label.metadata[label.metadata$region == "RM",]
CLR.CP3k.mat.RM = t(compositions::clr(t(CP3k.mat.RM+1)))

# PCA
sample.pca.CLR.CP3k.RM = prcomp(t(CLR.CP3k.mat.RM), center = TRUE, scale = TRUE)
pca.df.plotting = data.frame(sample.pca.CLR.CP3k.RM$x, timepoint = label.metadata.RM$timepoint)
ggplot(pca.df.plotting, aes(PC1, PC2, col = timepoint)) + geom_point(size = 3) + scale_color_manual(values = RM.palette.all[1:5])
```

In order to project into the existing PC space, we remove all unknown and unassigned cells and fix some data types in the dataframe
```{r}
rechallenge.md.annot = rechallenge@meta.data
rechallenge.md.annot$assignment = as.character(rechallenge.md.annot$assignment)

rechallenge.md.annot.known = rechallenge.md.annot[rechallenge.md.annot$predicted.id != "unknown",]
rechallenge.md.annot.known = rechallenge.md.annot[rechallenge.md.annot$assignment != "",]
```

Next we tabulate counts per cluster per assignment and calculate CP3k
```{r}
rechallenge.freq.table = rechallenge.md.annot.known %>% select(orig.ident, assignment, cell.type.cons, predicted.id, timepoint) %>% 
  group_by(assignment, predicted.id, .drop = FALSE) %>% summarise(n = n()) %>%
  mutate(CP3k = n/sum(n)*3e3) %>% ungroup() %>% 
  tidyr::complete(assignment, predicted.id, fill = list(n = 0, CP3k = 0))
```

create matrix for PCA
```{r}
CP3k.rechallenge = matrix(rechallenge.freq.table$CP3k, byrow = FALSE, nrow = length(unique(rechallenge.freq.table$predicted.id)), 
                             dimnames = list(unique(rechallenge.freq.table$predicted.id), unique(rechallenge.freq.table$assignment)))

#we have to add empty rows for those clusters not assigned
missed.clusters = rownames(CP3k.mat.RM)[!rownames(CP3k.mat.RM) %in% rownames(CP3k.rechallenge)]
not.assigned.mat = matrix(0, nrow = length(missed.clusters), ncol = ncol(CP3k.rechallenge), dimnames = list(missed.clusters))
CP3k.rechallenge = rbind(CP3k.rechallenge, not.assigned.mat)
CP3k.rechallenge = CP3k.rechallenge[rownames(CP3k.mat.RM),]
CLR.CP3k.rechallenge = t(compositions::clr(t(CP3k.rechallenge+1)))
```

project new matrix onto RM PCA
```{r}
# perform projection
predicted.X.pca = scale(t(CLR.CP3k.rechallenge), sample.pca.CP3k.RM$center, sample.pca.CP3k.RM$scale) %*% sample.pca.CP3k.RM$rotation
predicted.X.pca.df = data.frame(predicted.X.pca, timepoint = rep(c("D60_PR8","C2_PR8","C5_PR8","D60_X31","C2_X31","C5_X31"), times = 3)) #make sure the sample ordering is correct

# collapse the projection results with the original PCA emedding
pca.df.plotting2 = rbind(pca.df.plotting, predicted.X.pca.df)
pca.df.plotting2$experiment = c(rep("primary",15),rep("memory",18))
pca.df.plotting2$timepoint = factor(pca.df.plotting2$timepoint, levels = names(RM.palette.all))

ggplot(pca.df.plotting2, aes(PC1, PC2, col = timepoint, shape = experiment)) + geom_point(size = 3) + scale_color_manual(values = RM.palette.all)
ggsave("output/rechallenge_projected_PCA_CP3k.pdf", height = 5, width = 6)

ggplot(pca.df.plotting2, aes(PC3, PC4, col = timepoint, shape = experiment)) + geom_point(size = 3) + scale_color_manual(values = RM.palette.all)
ggsave("output/echallenge_projected_PCA_CP3k_PC3-4.pdf", height = 5, width = 6)
```


## Generate a new PCA using all data
Combine primary and memory abundance data
```{r}
CLR.CP3k.all = cbind(CLR.CP3k.mat.RM, CLR.CP3k.rechallenge) #CLR is done by sample, so we can simply combine the matrices

rechallenge.replicate.md = data.frame(timepoint = rep(c("D60_PR8","C2_PR8","C5_PR8","D60_X31","C2_X31","C5_X31"), times = 3),
                                      sample = rep(c("D60_PR8","C2_PR8","C5_PR8","D60_X31","C2_X31","C5_X31"), times = 3),
                                      row.names = colnames(CLR.CP3k.rechallenge))

replicate.md.all = rbind(label.metadata.RM[,-1], rechallenge.replicate.md)
replicate.md.all$experiment = c(rep("primary",15),rep("memory",18))
replicate.md.all$timepoint = factor(replicate.md.all$timepoint, levels = names(RM.palette.all))
```

Calculate and plot PCA
```{r}
all.replicates.PCA = prcomp(t(CLR.CP3k.all), center = TRUE, scale = TRUE)

pca.df.plotting.all = data.frame(all.replicates.PCA$x, timepoint = replicate.md.all$timepoint, experiment = replicate.md.all$experiment)

p1 = ggplot(pca.df.plotting.all, aes(PC1, PC2, col = timepoint, shape = experiment)) + geom_point(size = 3) + scale_color_manual(values = RM.palette.all)
p2 = ggplot(pca.df.plotting.all, aes(PC3, PC4, col = timepoint, shape = experiment)) + geom_point(size = 3) + scale_color_manual(values = RM.palette.all)
plot_grid(p1, p2, nrow = 1, rel_widths = c(1,1.02))
ggsave("output/primary_rechallenge_recalculated_PCA_1-4_CP3k.pdf", height = 5, width = 12)
```

What are the loadings?
```{r}
all.replicates.loadings = as.data.frame(all.replicates.PCA$rotation)
all.replicates.loadings$cluster = rownames(all.replicates.loadings)
write.xlsx(all.replicates.loadings, file = "output/primary_rechallenge_recalculated_PCA_loadings_CP3k.xlsx", rowNames=TRUE)

ggplot(as.data.frame(all.replicates.loadings), aes(y=PC1, x=reorder(cluster, PC1))) + geom_point(size = 0.5) + geom_hline(yintercept = 0, linetype = 2, col = "gray40") + 
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust = 1, size = 2)) + labs(x = "")
ggsave("output/primary_rechallenge_recalculated_PC1_Loadings.pdf", height = 4, width = 6)
ggplot(as.data.frame(all.replicates.loadings), aes(y=PC4, x=reorder(cluster, PC4))) + geom_point(size = 0.5) + geom_hline(yintercept = 0, linetype = 2, col = "gray40") + 
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust = 1, size = 2)) + labs(x = "")
ggsave("output/primary_rechallenge_recalculated_PC4_Loadings.pdf", height = 4, width = 6)
```

What are the distances between time points?
```{r}
aitchison = dist(t(CLR.CP3k.all), method = "euclidean", diag = TRUE, upper = TRUE)
aitchison.mat = as.matrix(aitchison)
aitchison.df = rbind(as.character(replicate.md.all$timepoint), as.data.frame(aitchison.mat))
write.xlsx(aitchison.df, file = "output/primary_rechallenge_aitchison_distance_mat.xlsx", rowNames=TRUE)

#get all pairwise distances
replicate.pairs = t(combn(colnames(aitchison.mat),2))
aitchison.pairs = data.frame(replicate.pairs, dist=aitchison.mat[replicate.pairs])
aitchison.pairs$pair = paste(sapply(strsplit(aitchison.pairs$X1, split="_"), "[[", 3),
                             sapply(strsplit(aitchison.pairs$X2, split="_"), "[[", 3),
                             sep = " - ")
write.xlsx(aitchison.pairs, file = "output/primary_rechallenge_aitchison_distance_pairs.xlsx", rowNames = FALSE)
```

** distance plots were generated in Prism **

# Abundance Analysis
## Split by cell type
Create joint metadata matrices
```{r}
cell.types = c("B Cell", "Endothelial", "Epithelial", "Fibroblast", "Granulocyte", "Myeloid", "Neuron", "Stromal", "T/NK Cell")
columns.to.keep = c("orig.ident","nCount_RNA","nFeature_RNA","assignment","percent.mt","timepoint","cell.type", "cluster.names")

# Split up the primary infection metadata by cell type
primary.md = RM@meta.data[,colums.to.keep]
primary.md$prediction.score.max = NA
primary.md$experiment = "primary"

primary.md.list = lapply(cell.types, function(x) {
  subset(RM, cell.type == x)
  })

# Prepare rechallenge metadata columns to merge appropriately
memory.md = rechallenge@meta.data[,c("orig.ident","nCount_RNA","nFeature_RNA","assignment","percent.mt","timepoint","cell.type.cons","predicted.id","prediction.score.max")]
colnames(memory.md) = c("orig.ident","nCount_RNA","nFeature_RNA","assignment","percent.mt","timepoint","cell.type","cluster.names","prediction.score.max")
memory.md$experiment = "memory"
memory.md$timepoint = as.character(memory.md$timepoint)

# combine cell type by cell type
combined.md.list = list()
for (i in 1:length(primary.md.list)){
  combined.md.list[[i]] = rbind(primary.md.list[[i]], memory.md[memory.md$cell.type == cell.types[i],])
  combined.md.list[[i]]$timepoint = factor(combined.md.list[[i]]$timepoint, levels = names(RM.palette.all))
}
names(combined.md.list) = names(cell.types)
```

Calculate the frequencies per cluster within each cell type (not used in the paper)
Note, I'm not specifying a filename to prevent plotting
```{r}
B.freq = CalcFreqByHashFromMD(metadata = combined.md.list[["B"]], 
                     col = cluster.names,
                     which.x.axis = timepoint,
                     palette = RM.palette.all,
                     filename = NULL,
                     plot.height = 12, plot.width = 12, plot.ncol = 4
                     )

endo.freq = CalcFreqByHashFromMD(metadata = combined.md.list[["endo"]], 
                     col = cluster.names,
                     which.x.axis = timepoint,
                     palette = RM.palette.all,
                     filename = NULL,
                     plot.height = 12, plot.width = 12, plot.ncol = 4
                     )

epi.freq = CalcFreqByHashFromMD(metadata = combined.md.list[["epi"]], 
                     col = cluster.names,
                     which.x.axis = timepoint,
                     palette = RM.palette.all,
                     filename = NULL,
                     plot.height = 24, plot.width = 12, plot.ncol = 4
                     )

fibro.freq = CalcFreqByHashFromMD(metadata = combined.md.list[["fibro"]], 
                     col = cluster.names,
                     which.x.axis = timepoint,
                     palette = RM.palette.all,
                     filename = NULL,
                     plot.height = 16, plot.width = 12, plot.ncol = 4
                     )

gran.freq = CalcFreqByHashFromMD(metadata = combined.md.list[["gran"]], 
                     col = cluster.names,
                     which.x.axis = timepoint,
                     palette = RM.palette.all,
                     filename = NULL,
                     plot.height = 12, plot.width = 12, plot.ncol = 4
                     )

mye.freq = CalcFreqByHashFromMD(metadata = combined.md.list[["mye"]], 
                     col = cluster.names,
                     which.x.axis = timepoint,
                     palette = RM.palette.all,
                     filename = NULL,
                     plot.height = 20, plot.width = 12, plot.ncol = 4
                     )

neuron.freq = CalcFreqByHashFromMD(metadata = combined.md.list[["neuron"]], 
                     col = cluster.names,
                     which.x.axis = timepoint,
                     palette = RM.palette.all,
                     filename = NULL,
                     plot.height = 12, plot.width = 12, plot.ncol = 4
                     )

stroma.freq = CalcFreqByHashFromMD(metadata = combined.md.list[["stroma"]], 
                     col = cluster.names,
                     which.x.axis = timepoint,
                     palette = RM.palette.all,
                     filename = NULL,
                     plot.height = 8, plot.width = 12, plot.ncol = 4
                     )

TNK.freq = CalcFreqByHashFromMD(metadata = combined.md.list[["TNK"]], 
                     col = cluster.names,
                     which.x.axis = timepoint,
                     palette = RM.palette.all,
                     filename = NULL,
                     plot.height = 16, plot.width = 12, plot.ncol = 4
                     )
```

## All cells
### Abundance by cell type per timepoint
Get the full metadata table
```{r}
names(combined.md.list) = NULL
combined.md.flat = do.call(rbind, combined.md.list)
```

Stacked bar chart of cell type by timepoint
```{r}
cell.type.palette.2 = cell.type.palette[-10] #note we remove HSCs

timepoint.by.cell.type = combined.md.flat %>% group_by(timepoint, cell.type) %>% summarise(count = n()) %>% mutate(freq = count/sum(count))
timepoint.by.cell.type$cell.type = factor(timepoint.by.cell.type$cell.type, levels = names(cell.type.palette))
ggplot(timepoint.by.cell.type, aes(x=timepoint, y = freq*100, fill=cell.type)) + geom_bar(stat="identity") + 
  scale_fill_manual(values = cell.type.palette) + theme(axis.text.x = element_text(angle = 45, hjust=1), axis.title.y=element_text(size = 16)) + 
  labs(x="", y="% of cells", fill = "Cell Type")
ggsave("output/timpoint_by_cell_type_barchart.pdf", height = 5, width = 6)
```

Replicate abundances by timepoint
```{r}
timepoint.by.hash = CalcFreqByHashFromMD(metadata = combined.md.flat, 
                     col = cell.type,
                     which.x.axis = timepoint,
                     palette = RM.palette.all,
                     filename = "output/cell_type_timepoint_by_hash_replicate.pdf",
                     plot.height = 10,
                     plot.width = 8,
                     plot.ncol = 3,
                     )
```

Specific plots to overlap the X31 and PR8 data on top of each other in time
```{r}
timepoint.by.hash$timepoint.grouped = plyr::mapvalues(timepoint.by.hash$timepoint, 
                                                      from = c("D60_PR8","C2_PR8","C5_PR8","D60_X31","C2_X31","C5_X31"), 
                                                      to = rep(c("D60","C2","C5"),2))

p.mye = ggplot(subset(timepoint.by.hash, Cluster.Names == "Myeloid"), aes(x = timepoint.grouped, y = freq*100, col = timepoint)) + 
  geom_jitter(width = 0.2, height = 0, size = 2.5) + 
  scale_color_manual(values = RM.palette.all) + labs(x = "", y = "% of RM cells", color = "Sample", title = "Myeloid") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + ylim(0, 50)
p.TNK = ggplot(subset(timepoint.by.hash, Cluster.Names == "T/NK Cell"), aes(x = timepoint.grouped, y = freq*100, col = timepoint)) + 
  geom_jitter(width = 0.2, height = 0, size = 2.5) + 
  scale_color_manual(values = RM.palette.all) + labs(x = "", y = "% of RM cells", color = "Sample", title = "T/NK Cells") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + ylim(0,45)
plot_grid(p.mye, p.TNK, ncol = 2)
ggsave("output/myeloid_and_TNK_cell.type_freqs_overlaid.pdf", height = 4, width = 9)

p.gran = ggplot(subset(timepoint.by.hash, Cluster.Names == "Granulocyte"), aes(x = timepoint.grouped, y = freq*100, col = timepoint)) + 
  geom_jitter(width = 0.2, height = 0, size = 2.5) + 
  scale_color_manual(values = RM.palette.all) + labs(x = "", y = "% of RM cells", color = "Sample", title = "Granulocyte") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + ylim(0, 15)
p.B = ggplot(subset(timepoint.by.hash, Cluster.Names == "B Cell"), aes(x = timepoint.grouped, y = freq*100, col = timepoint)) + 
  geom_jitter(width = 0.2, height = 0, size = 2.5) + 
  scale_color_manual(values = RM.palette.all) + labs(x = "", y = "% of RM cells", color = "Sample", title = "B Cells") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + ylim(0,35)
plot_grid(p.gran, p.B, ncol = 2)
ggsave("output/gran_and_B_cell.type_freqs_overlaid.pdf", height = 4, width = 9)
```


### Abundance by cluster per assignment (replicate)
Generate the frequency/abundance table and change into matrix
```{r}
allcells.freq =  CalcFreqByHashFromMD(metadata = combined.md.flat, 
                     col = cluster.names,
                     which.x.axis = timepoint)

allcells.freq = allcells.freq[-which(allcells.freq$Cluster.Names == "unknown"), ]

allcells.freq$CP3k = allcells.freq$freq*3e3
allcells.freq$log2CP3k = log2(allcells.freq$CP3k + 1) #used for plotting abundances

counts.mat = matrix(allcells.freq$count, byrow = FALSE, nrow = length(unique(allcells.freq$Cluster.Names)), 
                             dimnames = list(unique(allcells.freq$Cluster.Names), unique(allcells.freq$assignment)))
CP3k.mat = matrix(allcells.freq$CP3k, byrow = FALSE, nrow = length(unique(allcells.freq$Cluster.Names)), 
                             dimnames = list(unique(allcells.freq$Cluster.Names), unique(allcells.freq$assignment)))
CLR.CP3k.mat = t(compositions::clr(t(CP3k.mat+1)))

counts.df = rbind(unlist(lapply(names(RM.palette.all), rep, 3)), as.data.frame(counts.mat))
CP3k.df = rbind(unlist(lapply(names(RM.palette.all), rep, 3)), as.data.frame(CP3k.mat))
CLR.CP3k.df = rbind(unlist(lapply(names(RM.palette.all), rep, 3)), as.data.frame(CLR.CP3k.mat))
rownames(CP3k.df)[1] = rownames(counts.df)[1] = "Time Point"
```

<!-- write out the abundance matrix -->
<!-- ```{r} -->
<!-- write.xlsx(counts.df, file = "output/all_samples_primary_memory_abundance_matrix_counts.xlsx", rowNames = TRUE) -->
<!-- write.xlsx(CP3k.df, file = "output/all_samples_primary_memory_abundance_matrix_CP3k.xlsx", rowNames = TRUE) -->
<!-- ``` -->


### Comparing trajectories within cluster between primary & memory
First we need to add the experiment column to the allcells.freq table and add a new time column
```{r}
allcells.freq$experiment = plyr::mapvalues(allcells.freq$timepoint, from = names(RM.palette.all), 
                                           to = c(rep("primary", times = 5), rep("mem_PR8", times = 3), rep("mem_X31", times = 3)))
allcells.freq$timepoint.comp = plyr::mapvalues(allcells.freq$timepoint, from = names(RM.palette.all), 
                                               to = c("Baseline","2 dpc","5 dpc","8 dpc","14 dpc","Baseline","2 dpc","5 dpc", "Baseline", "2 dpc", "5 dpc"))
```

Plot the comparative graphs between Primary, X31, and PR8 rechallenge
```{r, warning=FALSE, message=FALSE}
cluster.list = c("IFN-Stim Mature Neu", "MHC-I Hi Mature Neu", "IFN-Stim MDM", "IFN-Stim Mono",
                 "Gzmk+ CD8 T", "Cd103+ CD8 T",  "Ifng+Cd200+ CD4 T", "CD4 T",
                 "Mature B 2", "IgG+/IgA+ B", 
                 "Krt13+Il1a+ Epi", "Meg3+MHC-II+ Epi", "Cycling Basal", "IFN-Stim Epi",
                 "MHC-II Hi Mac")
plot.list = list()
for(i in 1:length(cluster.list)){
  plot.list[[i]] = PlotTemporalAbundanceComparison(abundances = allcells.freq, cluster = cluster.list[[i]], 
                                                   group.palette = c("#007200", "#0065b8", "#9b14c4"), time.axis = timepoint.comp, span = 0.7) & NoLegend()
}

plot_grid(plotlist = plot.list, ncol = 4, byrow = TRUE)
ggsave("output/primary_vs_memory_abundances.pdf", height = 14.67, width = 10)
```


# Differential Gene Expression Analysis
## IFN-Stim MDMs
Subset and correct SCT normalization
```{r}
IFN_Stim_MDMs = merge(subset(RM, subset = cluster.names == "IFN-Stim MDM"), subset(rechallenge, subset = predicted.id == "IFN-Stim MDM"))
IFN_Stim_MDMs = subset(IFN_Stim_MDMs, subset = timepoint %in% c("D08","C2_PR8","C2_X31"))
IFN_Stim_MDMs$timepoint = factor(IFN_Stim_MDMs$timepoint, levels = c("D08","C2_PR8","C2_X31"))
IFN_Stim_MDMs = SetIdent(IFN_Stim_MDMs, value = "timepoint")
IFN_Stim_MDMs = PrepSCTFindMarkers(IFN_Stim_MDMs, assay = "SCT", verbose = TRUE)
```

Timepoint comparisons
```{r}
MDMs.all.markers = FindAllMarkers(IFN_Stim_MDMs, logfc.threshold = 0.5, min.pct = 0.2, only.pos = FALSE, test.use = "bimod")
write.xlsx(MDMs.all.markers, file = "output/IFN_Stim_MDMs_DEGs_v2.xlsx", rowNames = TRUE)
```

Plotting
```{r}
MDM.plot.markers = c("Lgmn","Cd72","Ccr1","Ms4a7","Rps29","Tnfaip2","Lars2","Parp14")
DotPlot(IFN_Stim_MDMs, features = MDM.plot.markers, scale = FALSE, dot.scale = 5, cols = c("purple4","gold"),) + coord_flip() + labs(x="", y="") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggsave("output/IFN-Stim_MDMs_DEGs_Dotplot_v2.pdf", height = 4, width = 4)
```

## CD103+ CD8 T C2
Subset
```{r}
CD103_CD8_T = merge(subset(RM, subset = cluster.names == "Cd103+ CD8 T"), subset(rechallenge, subset = predicted.id == "Cd103+ CD8 T"))
CD103_CD8_T = subset(CD103_CD8_T, subset = timepoint %in% c("D14","C2_PR8","C2_X31"))
CD103_CD8_T$timepoint = factor(CD103_CD8_T$timepoint, levels = c("D14","C2_PR8","C2_X31"))
CD103_CD8_T = SetIdent(CD103_CD8_T, value = "timepoint")
CD103_CD8_T = PrepSCTFindMarkers(CD103_CD8_T, assay = "SCT", verbose = TRUE)
```

Timepoint comparisons
```{r}
Trm.all.markers = FindAllMarkers(CD103_CD8_T, logfc.threshold = 0.5, min.pct = 0.2, only.pos = FALSE, test.use = "bimod")
write.xlsx(Trm.all.markers, file = "output/Cd103_CD8_T_DEGs_C2.xlsx")
```

Plotting
```{r}
Trm.plot.markers = c("Id2","Rgs1","Cxcr6","Ccr2","Rps29","Xcl1","Bcl2","Klf2")
#VlnPlot(CD103_CD8_T, features = Trm.plot.markers, cols = RM.palette.all, stack = TRUE, flip = TRUE)
DotPlot(CD103_CD8_T, features = Trm.plot.markers, scale = FALSE, dot.scale = 5, cols = c("purple4","gold")) + coord_flip() + labs(x="", y="") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggsave("output/Cd103_CD8_T_DEGs_Dotplot_C2.pdf", height = 4, width = 4)
```


<!-- ## CD103+ CD8 T C5 -->
<!-- Subset -->
<!-- ```{r} -->
<!-- CD103_CD8_T = merge(subset(RM, subset = cluster.names == "Cd103+ CD8 T"), subset(rechallenge, subset = predicted.id == "Cd103+ CD8 T")) -->
<!-- CD103_CD8_T = subset(CD103_CD8_T, subset = timepoint %in% c("D14","C5_PR8","C5_X31")) -->
<!-- CD103_CD8_T$timepoint = factor(CD103_CD8_T$timepoint, levels = c("D14","C5_PR8","C5_X31")) -->
<!-- CD103_CD8_T = SetIdent(CD103_CD8_T, value = "timepoint") -->
<!-- CD103_CD8_T = PrepSCTFindMarkers(CD103_CD8_T, assay = "SCT", verbose = TRUE) -->
<!-- ``` -->

<!-- Timepoint comparisons -->
<!-- ```{r} -->
<!-- Trm.all.markers = FindAllMarkers(CD103_CD8_T, logfc.threshold = 0.5, min.pct = 0.2, only.pos = FALSE, test.use = "bimod") -->
<!-- write.xlsx(Trm.all.markers, file = "output/Cd103_CD8_T_DEGs_C5.xlsx") -->
<!-- ``` -->

<!-- Plotting -->
<!-- ```{r} -->
<!-- Trm.plot.markers = c("Id2","Rgs1","Cxcr6","Ccr2","Rps29","Xcl1","Bcl2","Klf2") -->
<!-- #VlnPlot(CD103_CD8_T, features = Trm.plot.markers, cols = RM.palette.all, stack = TRUE, flip = TRUE) -->
<!-- DotPlot(CD103_CD8_T, features = Trm.plot.markers, scale = FALSE, dot.scale = 5, cols = c("purple4","gold")) + coord_flip() + labs(x="", y="") -->
<!-- ggsave("DEG_plots/Cd103_CD8_T_DEGs_Dotplot_C2.pdf", height = 4, width = 4.5) -->
<!-- ``` -->

## Gzmk+ CD8 T
Subset
```{r}
Gzmk_CD8_T = merge(subset(RM, subset = cluster.names == "Gzmk+ CD8 T"), subset(rechallenge, subset = predicted.id == "Gzmk+ CD8 T"))
Gzmk_CD8_T = subset(Gzmk_CD8_T, subset = timepoint %in% c("D08","C5_PR8","C5_X31"))
Gzmk_CD8_T$timepoint = factor(Gzmk_CD8_T$timepoint, levels = c("D08","C5_PR8","C5_X31"))
Gzmk_CD8_T = SetIdent(Gzmk_CD8_T, value = "timepoint")
Gzmk_CD8_T = PrepSCTFindMarkers(Gzmk_CD8_T, assay = "SCT", verbose = TRUE)
```

Timepoint comparisons
```{r}
Gzmk_CD8.markers = FindAllMarkers(Gzmk_CD8_T, logfc.threshold = 0.5, min.pct = 0.2, only.pos = FALSE, test.use = "bimod")
write.xlsx(Gzmk_CD8.markers, file = "output/Gzmk_CD8_T_DEGs.xlsx", rowNames = TRUE)
```

Plotting
```{r}
Gzmk_CD8.plot.markers = c("mt-Nd2","Icos","Ly6c2","Gzmb","Lag3","Fos","Selenoh","Tubb5","H2afv","Birc5")
DotPlot(Gzmk_CD8_T, features = Gzmk_CD8.plot.markers, scale = FALSE, dot.scale = 5, cols = c("purple4","gold")) + coord_flip() + labs(x="", y="") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggsave("output/Gzmk_CD8_T_DEGs_Dotplot.pdf", height = 4, width = 4)
```

## Ifng+Cd200+ CD4 T
Subset
```{r}
Ifng_CD4_T = merge(subset(RM, subset = cluster.names == "Ifng+Cd200+ CD4 T"), subset(rechallenge, subset = predicted.id == "Ifng+Cd200+ CD4 T"))
Ifng_CD4_T = subset(Ifng_CD4_T, subset = timepoint %in% c("D08","C2_PR8","C2_X31"))
Ifng_CD4_T$timepoint = factor(Ifng_CD4_T$timepoint, levels = c("D08","C2_PR8","C2_X31"))
Ifng_CD4_T = SetIdent(Ifng_CD4_T, value = "timepoint")
Ifng_CD4_T = PrepSCTFindMarkers(Ifng_CD4_T, assay = "SCT", verbose = TRUE)
```

Timepoint comparisons
```{r}
Ifng_CD4.markers = FindAllMarkers(Ifng_CD4_T, logfc.threshold = 0.5, min.pct = 0.2, only.pos = FALSE, test.use = "bimod")
write.xlsx(Ifng_CD4.markers, file = "output/Ifng_CD4_T_DEGs.xlsx", rowNames = TRUE)
```

Plotting
```{r}
Ifng_CD4.plot.markers = c("Cd200","Icos","Tnfrsf4","Batf","Itga1","Rps29","Batf3","Mt2")
DotPlot(Ifng_CD4_T, features = Ifng_CD4.plot.markers, scale = FALSE, dot.scale = 5, cols = c("purple4","gold")) + coord_flip() + labs(x="", y="") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggsave("output/Ifng_CD4_T_DEGs_Dotplot.pdf", height = 4, width = 4)
```

## Krt13+Il1a+ Epi (not reported)
Subset
```{r}
epikiller = merge(subset(RM, subset = cluster.names == "Krt13+Il1a+ Epi"), subset(rechallenge, subset = predicted.id == "Krt13+Il1a+ Epi"))
epikiller = subset(epikiller, subset = timepoint %in% c("D14","D60_PR8","D60_X31","C2_PR8","C2_X31","C5_PR8","C5_X31"))
epikiller$timepoint = factor(epikiller$timepoint, levels = names(RM.palette.all))
epikiller = SetIdent(epikiller, value = "timepoint")
epikiller = PrepSCTFindMarkers(epikiller, assay = "SCT", verbose = TRUE)
```

Timepoint comparisons
```{r}
epikiller.all.markers = FindAllMarkers(epikiller, logfc.threshold = 0.5, min.pct = 0.2, test.use = "bimod", only.pos = TRUE)
epikiller.all.markers = epikiller.all.markers[epikiller.all.markers$p_val_adj <= 0.05,]
write.xlsx(Trm.all.markers, file = "primary_and_memory/DEG_plots/epikiller_DEGs.xlsx")
```

Plotting
```{r}
epikiller.plot.markers = c("Ifit3","Irfd1","Calm1","Src","Cd81","Gm10076")
DotPlot(epikiller, features = epikiller.plot.markers, scale = FALSE, dot.scale = 5, cols = c("purple4","gold")) + coord_flip() + labs(x="", y="")
ggsave("primary_and_memory/DEG_plots/epikiller_DEGs_Dotplot.pdf", height = 2.5, width = 4.7)
```


# Combine RM and rechallenge
## Rerunning the UMAP
Rerun UMAP and save model
```{r}
RM = RunUMAP(RM, dims = 1:40, n.neighbors = 30, return.model = TRUE)
RM.anchors = RM.anchors = FindTransferAnchors(reference = RM,
                                 query = rechallenge,
                                 normalization.method = "SCT",
                                 reference.assay = "SCT",
                                 reference.reduction = "pca",
                                 dims = 1:40)
```

Project UMAP
```{r}
rechallenge.projected = IntegrateEmbeddings(anchorset = RM.anchors, reference = RM, query = rechallenge, new.reduction.name = "ref.pca")
rechallenge.projected = ProjectUMAP(query = rechallenge.projected, query.reduction = "ref.pca", reference = RM, reference.reduction = "pca", reduction.model = "umap")
```

## Renaming for combining
Create duplicate columns for cell type and cluster names so they can match between objects
```{r}
RM$cell.type.f = RM$cell.type; RM$cluster.name = RM$cluster.names
rechallenge.projected$cell.type.f = rechallenge.projected$cell.type.cons; rechallenge.projected$cluster.name = rechallenge.projected$predicted.id
```

Subset metadata matrices
```{r}
RM.md.cols.keep = c("orig.ident","nCount_RNA","nFeature_RNA","demux_type","assignment","nCount_HTO","nFeature_HTO",
                    "percent.mt","nCount_SCT","nFeature_SCT","timepoint","cell.type.f","cluster.name")
rechallenge.md.cols.keep = c(RM.md.cols.keep, "prediction.score.max")

RM@meta.data = RM@meta.data[,RM.md.cols.keep]
rechallenge.projected@meta.data = rechallenge.projected@meta.data[,rechallenge.md.cols.keep]
```

Merge
```{r}
rechallenge.projected$id <- 'reference'
RM$id <- 'query'
RM.PM.comb.2 <- merge(RM, rechallenge.projected)
RM.PM.comb.2[["umap"]] <- merge(RM[["umap"]], rechallenge.projected[["ref.umap"]])
```

Save
```{r}
save(RM.PM.comb.2, file = "output/RM_primary_memory_combined.RData", compress = FALSE)
```

